---
title: "Assignment3"
author: "Yi Dou"
format: 
  html:
    embed-resources: true
editor: visual
warning: false
message: false
bibliography: references.bib
---

## Exercise 03

```{r, message = FALSE}
library(tidyverse)
crime_data <- read_csv("data/Crime_Incidents_in_2025.csv")
crime_data <- crime_data |>
  select(CCN, REPORT_DAT, START_DATE, END_DATE, BLOCK, 
         OFFENSE, METHOD, SHIFT, LATITUDE, LONGITUDE) 
crime_data |>
  glimpse()
```

This project uses administrative crime data from Washington, DC (@dc_crime_2025). The time period represented by the data is from 2025-01-01 to 2026-01-01(UTC). The data were collected through administrative records from the Metropolitan Police Department based on reported crime incidents. Each observation represents a single reported crime incident and can be uniquely identified using the Crime Control Number (CCN). The data set is administrative and cross-sectional data including numeric, categorical, timeseries and geospatial variables. It does not include weights.

## Exercise 04

```{r}
sapply(crime_data, class)
crime_data <- crime_data |>
  mutate(
    report_datetime = ymd_hms(REPORT_DAT, tz = "UTC"),
    start_datetime  = ymd_hms(START_DATE,  tz = "UTC"),
    end_datetime    = ymd_hms(END_DATE,    tz = "UTC")
  )
str(crime_data$report_datetime)
```

I examined the variable class of the data set. While other variables aligned with the data documentation, the three datetime variables were stored as strings. Luckily, they had all the information I needed, including date, time, and time zone. So I used the function in the lubridate package to transform them into datetime format.
